`localhierarchy`: an R package to facilitate fitting of global and local Bayesian hierarchical models

### 2025-10-08 paper/paper.qmd
clean up export and naming of functions

  This article presents a practical introduction to the package for the applied user and illustrates the package’s functionality through examples for national and subnational estimation.
bibliography: bibliography.bib
number-sections: true
---

```{r}
#| include: false
library(here)
library(tidyverse)
library(ggplot2)
library(haven)
library(stringr)
library(cmdstanr)
library(truncnorm) # for inits
library(bayesplot) # diagnostic plots
library(posterior) # else error for as_draws_list
library(tidybayes)
library(patchwork)
devtools::load_all(here::here())
```

@@

![Schematic illustration of transition model used in the modeling of family planning indicators, modified from @susmann_flexible_2025. The transition model describes changes in an indicator (here $\eta_{c,t}$ in terms of three principal parameters: the rate of change, the level in a reference year (here labeled with $\Omega_c$), and the asymptotic level (the maximum attainable level for an indicator (here labeled with $\lambda_c$).](fig/transition.png){#fig-transition width="476"}

@@

```{r}
plot_prior_post_sigmas(fit = global_fit, parname = "mu")
```

@@

>>>>

  This article presents a practical introduction to the package for the applied user and illustrates the package’s functionality through examples for national and subnational estimation.

keywords: Bayesian hierarchical model; R; Stan; Local models 
bibliography: bibliography.bib
number-sections: true
---

```{r}
#| include: false
library(here)
library(tidyverse)
library(ggplot2)
library(haven)
library(stringr)
library(cmdstanr)
library(truncnorm) # for inits
library(bayesplot) # diagnostic plots
library(posterior) # else error for as_draws_list
library(tidybayes)
library(patchwork)
library(localhierarchy)
```

@@

![Schematic illustration of transition model used in the modeling of family planning indicators. The transition model describes changes in an indicator (here $\eta_{c,t}$ in terms of three principal parameters: the rate of change, the level in a reference year (here labeled with $\Omega_c$), and the asymptotic level (the maximum attainable level for an indicator (here labeled with $\lambda_c$). This figure has been reproduced with permission from @susmann_flexible_2025.](fig/transition.png){#fig-transition width="476"}

@@

```{r}
plot_prior_post_sigmas_localhierarchy(fit = global_fit, parname = "mu")
```

@@

### Data availability statement {.unnumbered}

No data are associated with the software tool described in this article.

### Software availability {.unnumbered}

The R package is available at [github.com/AlkemaLab/localhierarchy](https://github.com/AlkemaLab/localhierarchy) and Zenodo \[to do: add reference\]. The code is licensed XXX.

### Grant information {.unnumbered}

This work was supported, in whole or in part, by the Gates Foundation (INV-00844 and OPP1192802). Under the grant conditions of the Foundation, a Creative Commons Attribution 4.0 Generic License has already been assigned to the Author Accepted Manuscript version that might arise from this submission.
